### 2021-11-30 items_english.rmd

---
title: "Items: English"
output:
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
      toc_depth: 3
---

Here you can find a quick overview of all items that were measured.

## Life Satisfaction

"Taken everything together, how satisfied are you currently with your life?"

(0 = extremely unsatisfied, 10 = extremely satisfied)

## Positive Affect (scale)

"In the last week, how often did you feel"

 - a: ... calm and relaxed
 - e: ... happy
 - i: ... full of energy

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

## Negative Affect (scale)

"In the last week, how often did you feel"

 - b: ... lonely
 - c: ... aggravated
 - d: ... so depressed, that nothing could lift you up
 - f: ... very nervous
 - g: ... anxious
 - h: ... glum and sad

(never [1]; on some days [2]; several times per week [3]; almost every day [4]; daily [5])

## Social media use

"How often during the last week have you engaged in the following activities on social media?"

(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)

### Reading

 - "Reading the posts of others with content on the Coronavirus."

### Liking & Sharing

 - "When seeing Posts on the Coronavirus, I clicked "like", "share" or "retweet".

### Active Postings

 - "I myself wrote posts on the Coronavirus on Social Media."

### Specific platforms

"How often in the last week have you followed information related to the Corona-crisis in the following social media?"

 - Facebook 
 - Twitter
 - Instagram
 - Youtube
 - WhatsApp

(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)

## Gender

"What is your biological sex?"

(a: Male; b Female; c: Diverse)

## Age

"In which year were you born?"

(year)

## Education

"What is the highest educational degree you have earned?"

 1. No degree
 1. Middle school
 1. Vocational school
 1. Technical school
 1. High school
 1. Applied high school
 1. State college
 1. Bachelor
 1. Master
 1. PhD

## Austria country of birth

"Is Austria you country of birth?"

  - Yes
  - NO

## Austria country of birth of parents

 "Is Austria your parents' country of birth?"

 - a. Yes, one parent.
 - b. Yes, both parents.
 - c. No.

## News consumption: Text-based

"Below you can find a list of different newspapers. How often during the last week have you informed yourself politically using on the these outlets (both online and offline)?"

 - Kronen Zeitung or www.krone.at
 - Der Standard or derstandard.at
 - Die Presse or diepresse.com
 - Österreich or oe24.at
 - Kurier or kurier.at
 - Salzburger Nachrichten or salzburg.at
 - Other Austrian papers

## News consumption: Video-based

"Below you can find a list of different news shows in television. How often during the last week have you informed yourself politically using on the these shows?"

 - "News shows on ORF (e.g., Zeit im Bild at 19:30h or ZiB 2 at 22:00h)"
 - "News shows in private television (e.g., ATV aktuell, Puls 4/ Pro7/ Sat1 Austria)"

(1 = several times per day; 2 = daily; 3 = several times per week; 4 = weekly; 5 = never)

## Residency Vienna

"In which county do you live currently?"

 - a. Vorarlberg
 - b. Tirol
 - c. Salzburg
 - d. Steiermark
 - e. Kärnten
 - f. Oberösterreich
 - g. Niederösterreich
 - h. Wien
 - i. Burgenland

(Was transformed to Vienna: yes/no)

## Household size

"Please include yourself into the following calculation. If one groups not present in your household, indicate "0"."

 - a. ...Adults older than 18 years:
 - b. ...Teenager between 15-17 years:
 - c. ...Kids between 6-14 years:
 - d. ...Kids between 3-5 years:
 - e. ...Kids between 0-2 years:

## Health

"Let's next turn to some questions concerning health. How would you describe your overall health conditionas? Would you say your healt is ..."

 - a. 1 = Sehr gut
 - b. 2 = Gut
 - c. 3 = Mittelmäßig
 - d. 4 = Schlecht
 - e. 5 = Sehr schlecht

## Living space

"How many square meteres does your appartment have?"

(Number, min. 0 – max. 600, voluntary.)

## Access garden

"Have you got access to a private area outside?"

 - (Yes, garden)

## Access balcony

"Have you got access to a private area outside?"

 - (Yes, balcony)

## Employment

In Februrary 2020, what was your job situation?

 1. Unemployed
 1. Industrie
 1. Public service
 1. Self-employed
 1. Retired
 1. Housekeeping
 1. Student
 1. Incapacitated
 1. Parental Leave

## Work hours per week

"How many hours per week (including extra hours) do you currently work in your main job? If you don't know, please guess.

(hours)

## Home-Office Yes/No

"Has your current job situation changed due to the Corona crisis? Please select the appropriate answer."

 - a. I'm in home-office

## Household income

"How much income does your household currently have per month?"

 - a. less than 1.100 Euro
 - b. between 1.100 and 1.500 Euro 
 - c. between 1.500 and 1.800 Euro 
 - d. between 1.800 and 2.200 Euro 
 - e. between 2.200 and 2.700 Euro 
 - f. between 2.700 and 3.100 Euro 
 - g. between 3.100 and 3.700 Euro 
 - h. between 3.700 and 4.300 Euro 
 - i. between 4.300 and 5.500 Euro 
 - j. 5.500 Euro and more

## Outdoor activities

"Have you left your home during the last week for the following activity?"

 - a: Work
 - b: Sport
 - c: Meeting friends or relatives
 - e: Shopping
 - f: Pets

(nie [1]; an manchen Tagen [2]; mehrmals die Woche [3]; beinahe jeden Tag [4]; täglich [5])

## Satisfaction Democracy

"How satisfied or dissatisfied are you all in all with the democracy how it's currently working in Austria?"

 - a. 1 = very satisfied
 - b. 2 = rather satisfied
 - c. 3 = neither nor
 - d. 4 = rather dissatisfied 
 - e. 5 = very dissatisfied

## Risk-taking

"How willing are you, in general, to take risks?"

(0: not at all willing; 10: very willing)

## Locus of Control

 - "I'm in control of my own life."
 - "If I try hard, I'll have success."
 - "Bei it in my private or professional life: My life is primarily determined by others."
 - "My plans for life are determined by fate."

(1 = Agree completely; 2 = Agree somewhat; 3 = Disgagree somewhat; 4 = Disagree completely)